---
title: 'Experiment3: More Airline Data'
subtitle: 
date: "2023-09-13"
output:
  html_document:
    df_print: paged
---

# Packages

```{r}
library(tidyverse)
library(nycflights13)
library(here)
library(janitor)
library(ggplot2)

```

# Get data from the online source.

Save the download to your data_raw directory. It will be a zipfile,, but that will be OK. 
Don't change the name of the zipfile. 

## Load the BOT data 

We take advantage of the 'here' package to get the path to the file that we stored. We also may note that inside the zipfile is a csv file. We will read the csvfile. 

```{r}
thisfile <- here("data_raw", "DL_SelectFields.zip")

df2022 <- read_csv(thisfile) |> clean_names()
```
## Subsetting to data of interesst

Load in all BOT data needed
```{r}
thisfile <- here("data_raw",  "DL_SelectFields_2017.zip")
df2017 <- read_csv(thisfile) |> clean_names()

thisfile <- here("data_raw",  "DL_SelectFields_2018.zip")
df2018 <- read_csv(thisfile) |> clean_names()

thisfile <- here("data_raw",  "DL_SelectFields_2019.zip")
df2019 <- read_csv(thisfile) |> clean_names()

thisfile <- here("data_raw",  "DL_SelectFields_2020.zip")
df2020 <- read_csv(thisfile) |> clean_names()

thisfile <- here("data_raw",  "DL_SelectFields_2021.zip")
df2021 <- read_csv(thisfile) |> clean_names()

thisfile <- here("data_raw", "DL_SelectFields_2022.zip")
df2022 <- read_csv(thisfile) |> clean_names()

thisfile <- here("data_raw", "DL_SelectFields_2023.zip")
df2023 <- read_csv(thisfile) |> clean_names()

```

Bind all data from 201-2023 together
```{r}
dfc=df2017 |>
  bind_rows(df2018)|>
  bind_rows(df2019)|>
  bind_rows(df2020)|>
  bind_rows(df2021)|>
  bind_rows(df2022)|>
  bind_rows(df2023)

```

```{r}
dfLGA = dfc |>
  filter(origin == "LGA" | dest == "LGA", passengers > 0)

dfLGA |>
  ggplot(aes(month)) +
  geom_bar(aes(weight = passengers)) 
```

```{r}
dfLGA |>
  ggplot(aes(year)) +
  geom_bar(aes(weight = passengers), fill = "red")
```
Drop in passengers at beginning of year (i.e January, February, March)
Most recent month has the most passengers out of all the months over all the years. 

# Passengers by month from 2017 to 2023
```{r}
dfLGA |> 
  mutate(time_year = year + month / 12) |>
  ggplot(aes(x = time_year)) + geom_bar(aes(weight = passengers), fill = "purple") 

```

```{r}

dfLGA_MIA = dfLGA |>
  mutate(by_month = ym(str_c(year,"-",month))) |> 
  #ggplot(aes(x = by_month)) + geom_bar(aes(weight = passengers), fill = "purple") 
  filter(dest == "MIA" | origin == "MIA")|>
  group_by(origin, by_month) |>
  summarise(passengers = sum(passengers)) |>
  mutate(cum_pass = cumsum(passengers)) |> 
  pivot_wider(names_from = origin, values_from = c("passengers", "cum_pass")) |>
  mutate(net_to_miami = passengers_LGA - passengers_MIA, cum_to_miami = cum_pass_LGA - cum_pass_MIA)

```
Going from all of NY to all of FL
`
```{r}
3 + 5
```

```{r}
dfLGA_MIA = dfLGA |>
  filter((origin == "MIA" | dest == "MIA") , passengers > 0) |>
  mutate(time_year = year + month / 12)

dfLGA_MIA |>
  group_by(origin, month) |>
  summarise(total_passengers = sum(passengers)) |>
  ggplot(aes(x = month, y = total_passengers)) + geom_line()

```

```{r}
dfLGA |>
  ggplot()
```

